---
title: "Lab 02 R Basics"
author: "Dr. Xiang Ji @ Tulane University"
date: "Sept 1, 2023"
output:
  html_document:
    toc: true
    toc_depth: 4  
subtitle: MATH-7360 Data Analysis
csl: ../apa.csl
---

## Announcement

- Remember to accept GitHub organization invites.

## Install Git

Now it's time that we try out Git, the (probably most) popular version control software.
Follow instructions on [installing Git](https://happygitwithr.com/install-git.html) from Happy Git and GitHub for the useR.
 
 - For windows users, I highly recommend you using the **Option 1** so that we are on the same page.
 
 - For mac users, open a terminal window and run command "xcode-select --install" that installs Xcode command line tools with Git for you.

## Git-related Yuja videos

Please watch the Yuja recordings on Canvas under `misc` of this week.
You want to learn how to create a Git repository under our GitHub organization.
You also want to learn how to submit your homework.
These videos are from 2020, don't forget to correct the course organization name.

### Get started with homework

On [GitHub.com](https://github.com):

- Obtain [Student Developer Pack](https://education.github.com/pack).  (Should be done by now)

- Create a **private** repository `Math-7360-FirstName-LastName` on `tulane-math-7360-2023`.

On your local machine: 

- Clone the repository, create `develop` branch, where you work on solutions.

    ```{bash, eval=FALSE}
    # clone the project
    git clone https://github.com/tulane-math-7360-2023/Math-7360-FirstName-LastName.git
    # enter project folder
    cd Math-7360-FirstName-LastName
    # what branches are there?
    git branch
    # create develop branch
    git branch develop
    # switch to the develop branch
    git checkout develop
    # create folder for HW1
    mkdir hw1
    cd hw1
    # let's write solutions
    echo "sample solution" > hw1.Rmd
    echo "some bug" >> hw1.Rmd
    # commit the code
    git add hw1.Rmd
    git commit -m "start working on problem #1"
    # push to remote repo
    git push
    ```


--------------------------------------------------------------------------------

## More on types

R has 5 basic or "atomic" classes of objects:

 - character

 - numeric (real numbers)
 
 - integer
 
 - complex (not used very often)
 
 - logical (True/False)
 
When working with _vectors_ in R, it is important to remember that a vector can only be composed of one data type.

What happens when you put multiple objects with different type to form a vector?

The answer is **coercion** (in other words, type conversion).

Coercion means changing the type of an object and it happens by a pre-set order:

logical < integer < numeric < character

When you create a vector from objects with different types, the lower-order ones will be coerced to the highest type by R.

```{r}
a <- c(1L, "This is a character")

b <- c(TRUE, "Hello World")

c <- c(FALSE, 2)  # what is wrong here?
```

> **Question 1** What types do the above vectors hold?


## Explicit coercion

Of course you can coerce objects from one class (type) to another.
We can do this by using "as.\*" functions where \* is the class (type) you want to coerce the object into.

```{r}
# using the same objects a, b, c from the above question
a.logical <- as.logical(a)
a.integer <- as.integer(a)
a.numeric <- as.numeric(a)

b.logical <- as.logical(b)
b.integer <- as.integer(b)
b.numeric <- as.numeric(b)

c.logical <- as.logical(c)
c.integer <- as.integer(c)
c.numeric <- as.numeric(c)
c.character <- as.character(c)


d <- -5:5
d.logical <- as.logical(d)
```

> **Question 2** What do you get after the coercions?  Does any one suprise you?  Can you figure out why?

Hint: what are the types of each element in the vectors before you explicitly coerce them?

## Try arithmetics

OK, now let's try arithmetics.

Easy, right?

First let's create a vector $\mathbf{v} = (969, 971, 972, \dots, 1022, 1023)$ of 54 elements. (Attention, there is no $970$ in the vector)

```{r}
# finish the code below
v <- c()
```

Then, let's compute the sum $\sum_{i=1}^{54}2^{v_i}$.

```{r}
# finish the code below
v.power.sum <- sum()
```

How about only sum over 53 elements $\sum_{i=2}^{54}2^{v_i}$  (note that the sum starts from $v_2$).

```{r}
# finish the code below
v.power.sum.53 <- sum()
```

Now let's try putting the first element back

```{r}
v.power.sum.second <- 2^v[1] + sum()
```

> **Question 3** Explain what you found.

