---
title: "Lab 09 binomial regression"
author: "Dr. Xiang Ji @ Tulane University"
date: "Nov 3, 2023"
output:
  html_document:
    toc: true
    toc_depth: 4  
subtitle: MATH-7360 Data Analysis
csl: ../apa.csl
---

```{r}
rm(list = ls()) # clean-up workspace
library(tidyverse)
library(faraway)
```

# Presentation order

```{r, warning = FALSE}
set.seed(7360)

student.groups <- matrix(c("Agbomola, Oluwasegun Joshua", "Zimmer, Mattie",
            "Vaduthala, Nathaniel", "Collopy, John",
            "Berko, Abena", "Weaver, Kathleen",
            "Alan, Baris", "Argentino, John",
            "Qu, Jiazheng", "Huang, Yongtai",
            "Mehran, Syed", "Olson, Aidan",
            "Kalgi, Ketan Vinod", "Zhang, Yipeng",
            "Carlino, Delia", "Zhu, Kyra",
            "Islam, Rubaiyat Bin", "de la Pena, Andrew",
            "Lopez Santander, John Jairo", "Trinh, Lan",
            "Uddin, Moslem", "Sakran, Naufil"), 11, 2, TRUE) %>%
  apply(1, paste, collapse = " & ") %>%
  sample() %>%
  matrix(3, 4, byrow = TRUE, dimnames = list(c("1st", "2nd", "3rd"),
                                            c("Dec 1", "Dec 4", "Dec 6", "Dec 8"))) 
student.groups[3, 4] <- ""
print(student.groups)
```


# Exercise 3.1 from ELMR.

The question concerns data from a case-control study of esophageal cancer in Ileet-Vilaine, France.
The data is distributed with R and may be obtained along with a description of the variables by:

```{r, eval = FALSE}
data(esoph)
help(esoph)
```

## (a) Plot the proportion of cases agianst each predictor using the size of the point to indicate the number of subject as seen in figure below.
Comment on the relationships seen in the plots.
```{r}
lmod <- glm(chd ~ height + cigs, family = binomial, wcgs)
gdf <- wcgs %>%
  mutate(residuals = residuals(lmod), linpred = predict(lmod)) %>%
  group_by(cigs) %>%
  summarise(residuals = mean(residuals), count = n()) 
gdf %>%
  ggplot(mapping = aes(x = cigs, y = residuals, size = sqrt(count))) +
  geom_point() +
  theme_bw()
```


## (b) Fit a binomial GLM with interactions between all three preditors.
Use AIC as a criterion to select a model using the `step` function.
Which model is selected?


## (c) 
All three factors are ordered and so special contrasts have been used approriate for ordered factors involving linear, quadratic and cubic terms.
Further simplification of the model may be possible by eliminating some of these terms.
Use the `unclass` function to convert the factors to a numerical representation and check whether the model may be simplified.


## (d)
Does your final model fit the data? Is the test you make accurate for this data?


## (e)
Check for outlier in your final model.

## (f)
What is the predicted effect of moving one category higher in alcohol consumption?


## (g)
Compute a 95% confidence interval for this predicted effect.

